### 2016-05-21 paper/mathematical_formulation.tex
added new figures (not yet made pretty) corresponding to new error breakdown for cd vs cdr case; added cost comparison (linear solve comparison)
Small edits, to be continued

%------------------------------------------------------------------------------%
\section{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%

We seek to infer a parameter $q \in Q$, where $Q$ is a Hilbert space. Consider a state variable $u \in U$, which satisfies a bilinear form $a:U \times U \to \Reals$, such that,
\begin{equation}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\label{eq:weakForm}
\end{equation}
where the form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses.

We define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations; we denote the actual observations by $d\in\R^{n_d}$. The unknown parameters can then be inferred by minimizing the difference between the predicted and actual observations. This inverse problem is often ill-posed; the observations are noisy and sparse and thus insufficiently informative to uniquely determine the parameters. To remedy this, regularization is used to inject prior information or beliefs about the parameters. The inverse problem with regularization can be written as a constrained optimization problem
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q) \\ \textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\end{array}
\label{eq:invOpt}
\end{equation}
where we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization penalty term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the model in Equation (\ref{eq:weakForm}).

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI.

In the following derivation, we assume $a$ is three times continuously differentiable with respect to the state $u$ and parameters $q$, $C$ is three times continuously differentiable with respect to the state $u$, $R$ is differentiable with respect to the parameters $q$, and $I$ is differentiable with respect to the state $u$ and parameters $q$.

%------------------------------------------------------------------------------%
\section[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse \\Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%

A given physical system need not have a unique model that can describe it; there may be various models of different fidelities. For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model, as compared to that which would have resulted from solving the inverse problem with the highest-fidelity model.
%
\begin{theorem}
\label{thm:error_estimate}
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\label{eq:finErrExp}
\end{multline}
\end{theorem}
%

@@

>>>>

%------------------------------------------------------------------------------%
\section{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
We seek to infer an unknown parameter $q \in Q$, where $Q$ is a Hilbert space. Consider a state variable $u \in U$, which satisfies a bilinear form $a:U \times U \to \Reals$, such that,
%
\begin{equation}
\label{eq:weakForm}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\end{equation}
%
where the form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses.

We define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$. The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation.

The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q) \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
where we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by Equation (\ref{eq:weakForm}). The constraints given by Eq.~\eqref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models, introducing a tradeoff between fidelity and computational expense.

In the case of a goal-oriented inverse problem, the ultimate purpose of inferring the unknown parameters is to calculate some Quantity of Interest (QoI). Assuming a single scalar QoI, we denote this QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. Since our ultimate goal is to compute this QoI, the particular tradeoff we consider is the fidelity and resultant computational expense of the model we use, versus the error in the QoI. We thus seek to introduce the QoI functional $I$ into the inverse problem formulation, by introducing auxilary variables and additional adjoint equations. Then we use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those which occur due to the use of constraints other than those given by Eq.~\eqref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\section[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse \\Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model, as compared to that which would have resulted from solving the inverse problem with the highest-fidelity model. The main result is,
%
\begin{theorem}
\label{thm:error_estimate}
In the following derivation, we assume $a$ is three times continuously differentiable with respect to the state $u$ and parameters $q$, $C$ is three times continuously differentiable with respect to the state $u$, $R$ is differentiable with respect to the parameters $q$, and $I$ is differentiable with respect to the state $u$ and parameters $q$.

\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\label{eq:finErrExp}
\end{multline}
\end{theorem}
%

@@

\red{We should include something about how to map error to elements to refine...}
